---
title: "Social Vulnerability Mapping"
author: "Randy"
date: '2022-06-23'
output: html_document
---

# Goals

* Make proof of concept maps combining the Wildfire Risk to Important Watersheds and SVI metrics for North Carolina
* Learn to wrangle data, and make the maps


* IMPORTANTE!  we need to "normalize" that is use rates if that is not already the case in the SVI data!  


## Install and load libraries

```{r setup, message= FALSE, warning= FALSE, include= FALSE}
#install.packages("tidyverse")
#install.packages("tigris")
#install.packages("usmap")
library(tidyverse) # data reading and wrangling
library(tigris) # for counties data
library(scales) # for commas (and other formatting) with charts
library(usmap)
library(sf)
```


## Read in then join datasets for SVI exploration


```{r data, message=FALSE, warning=FALSE}
# SVI master dataset
svi_data <- read_csv("data/SVI2018_US_COUNTY.csv")

wfp_data <- read_csv("data/countyMeanWFP.csv")

# counties data (has geometries).  Need to remove AK, HI, all STATEFP >57
us_cntys <- counties() %>%
  subset(STATEFP < 57) %>% #insular areas
  subset(STATEFP != "02")%>%  #Alaska
  subset(STATEFP != "15" )   #Hawaii
```

## Join svi to counties.  Note-I did not create a new dataframe, and only joined SVI metrics of interest.  You can easily change the selected for columns
```{r}

# do the join of svi to counties
cntys_svi_metrics <- left_join(us_cntys,
                              svi_data %>% dplyr::select(FIPS, E_POV, E_MINRTY, E_DISABL),
                               by = c("GEOID" = "FIPS"))

# remove unneeded columns

cntys_svi_metrics <- cntys_svi_metrics %>%
                      select(-c(7:15))

cntys_svi_wfp_metrics <- left_join(cntys_svi_metrics,
                                   wfp_data %>% dplyr::select(mean, GEOID),
                                   by = "GEOID"
                                   
)


# save as shapefile for mapping in QGIS

st_as_sf(cntys_svi_wfp_metrics)

st_write(cntys_svi_wfp_metrics, 
         "outputs/cntys_svi_wfp_metrics.shp")


```


## density charts

I only did one...you can do the rest.  

```{r}
e_pov_hist <-
  cntys_svi_metrics %>%
  ggplot(aes(x = E_POV)) +
  geom_histogram(fill = 'skyblue', color = 'grey23' ) +
  labs(
    title = "Density chart for E_POV metric",
    x = "You need to figure out the units for this :)"  
    )  +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma)

e_pov_hist
```



